Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-11-26 manuscript.qmd
remove remaining comments, make stuff dynamic

<!-- Over the course of this work, we will first briefly recap the HMM in @sec-unilateral, which was so far used to model ipsilateral lymphatic progression only. Then, we intuitively extend it to include the contralateral side as well in @sec-ext-to-contra. In this section, we also introduce a way of modelling the tumor's midline extension as a random variable (@sec-midline) and lastly talk about how it may affect the contralateral spread in @sec-params-symmetry. -->

<!-- TODO: Hint at methods and results... -->

@@

<!--
Our first model to predict the lymphatic progression of HNSCC was introduced using Bayesian networks [@pouymayou_bayesian_2019]. We subsequently extended this work to a hidden Markov model (HMM) [@ludwig_hidden_2021] to allow an intuitive inclusion of T-category into the predictions. We will briefly summarize this HMM's formalism before building on it to include the contralateral spread in @sec-ext-to-contra.
-->

@@

- A patient with a lateralized early T-category tumor and ipsilateral LNL II involvement has a predicted risk of 1–2% <!-- TODO: make this dynamic --> for occult contralateral LNL II disease (green histogram).  
- If the same tumor extends over the midline, the risk increases to {{< var risk.early_ext_iN0_cN0_II >}}% (orange histogram).

@@

- For an advanced T-category tumor extending over the midline with a clinically negative contralateral LNL II, the risk is only 2% <!-- TODO: make this dynamic --> (green histogram).  
- If contralateral LNL II is clinically involved, the risk for contralateral LNL III rises to 4–5% <!-- TODO: make this dynamic --> (blue histogram).  
- When contralateral LNL II involvement is confirmed by FNA, the risk further increases to {{< var risk.late_ext_iII_cII_FNA_III >}}% (red histogram).

@@

For contralateral LNL IV, the predicted risk is below 1% <!-- TODO: make this dynamic --> in most scenarios, even for advanced T-category tumors extending over the midline with extensive ipsilateral and contralateral involvement (green histogram).

>>>>

@@

@@

- A patient with a lateralized early T-category tumor and ipsilateral LNL II involvement has a predicted risk of {{< var risk.early_noext_iII_cN0_II >}}% for occult contralateral LNL II disease (green histogram).  
- If the same tumor extends over the midline, the risk increases to {{< var risk.early_ext_iN0_cN0_II >}}% (orange histogram).

@@

- For an advanced T-category tumor extending over the midline and with extensive clinical involvement in the ipsilateral LNLs II, III, and IV, but a clinically negative contralateral neck, the risk is only {{< var risk.late_ext_iII+III+IV_cN0_III >}}% (green histogram).  
- If contralateral LNL II is clinically involved, the risk for contralateral LNL III rises to {{< var risk.late_ext_iII_cII_III >}}% (blue histogram).  
- When contralateral LNL II involvement is confirmed by FNA, the risk further increases to {{< var risk.late_ext_iII_cII_FNA_III >}}% (red histogram).

@@

For contralateral LNL IV, the predicted risk is below 1% in most scenarios, even for advanced T-category tumors extending over the midline with extensive ipsilateral and contralateral involvement (green histogram).
